\section{Experiments}
\label{section:experiments}

To support our position that further work is needed to enable LLMs to reason spatially, we first design a set of experiments (E1 - E5) to illustrate the present limitations of LLMs with regard to spatial reasoning.
The first set of experiments probes spatial reasoning ability across targeted spatial entity types, and the second set targets reasoning over various complex spatial relations.


\subsection{Spatial Entities} % -----------------------------------------------

\subsubsection{E1. Non-point Data}
\cite{Liu2023} can do NL2Spatial Query which can handle region/line data, but can an LLM handle it?
--> Show this info isn't learned in LLMs by repeating queries similar to \cite{Liu2023} nanjingtest and berlintest region and line queries but instead give them to ChatGPT.

\subsubsection{E2. Lesser-known Cities}
To determine if LLMs are able to answer spatial questions about less populous locations and cities, we prompt Chat-GPT with the following question, filling in the blank with the name of a location in Australia.
The locations are selected based on their population size, ranging from \nrscomment{fill in details}
For each prompt, we measure whether the location was recognized or not \nrscomment{give unrecognized output phrase}.
We also measure whether the response is spatially accurate \nrscomment{decide if this is feasible}





\subsection{Spatial Relations} % -----------------------------------------------

\subsubsection{E3. N-way Spatial Relations}
--> Repeat AUS cities test from \cite{Qi2023} paper but with 3 way relations to show it isn't really doing spatial reasoning

\subsubsection{E4. Non-directional Spatial Relations}
Cyclic? Topological?
--> Show this info isn't learned in LLMs by making queries similar to the \cite{Qi2023} ones but for relations types besides directional.

\subsubsection{E5. Spatiotemporal and Multiple Hop Relations}
--> Test if Chat-GPT can answer queries like
- whether event x happened north of event y (2 hops event->loc + loc->spatial)
- all events in x region that happened between y and z dates (intersect space and time)
- all events in x region that happened between y and z dates north of location A (spatiotemporal involving spatial relation)








        
% Future Work -------------------------------------

% \subsection{SPM}
% \nrscomment{Experiment.}
% --> Try giving Chat-GPT a bunch of points A, B, C with coords like the pictorial query grid has and then asking it which cities in a given region match that pattern.
% Also try giving the input instead as a list of pairwise constraints.
% Use OSM to figure out the ground truth by pulling all city tags in the same region and running a traditional SPM algorithm to find all matching patterns.
% Report precision and recall for both input types.

% Check - can it produce an image from the points given? Can we give it an image with points as input?



% \paragraph{Locations being too close to differentiate in the embedding space}
% <ref from translation clustering paper>
% \nrscomment{Experiment.}
% --> Design tests to demonstrate these issues - similar to above, compare the embeddings of locations.